---
title: "FieldSurveyTutorial"
author: "Jasmine Nasser"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package analyzes and manipulates field survey data.

```{r}
#install packages for better access to the functions
devtools::install_github("jasminenasser/fieldsurvey.pkg")
#install.packages("tidyverse")
#install.packages("ggplot2")
library(tidyverse)
library(ggplot2)
library(nasser.pkg)
```

In this tutorial, I will use previous field data from 2005-2013 from North Dakota.

```{r}
#install.packages("googlesheets4")
#install.packages("googlesheets4")
library(googlesheets4)
googlesheets4::gs4_deauth()
AGP_data <- read_sheet("https://docs.google.com/spreadsheets/d/1hH39exibZwdQyr3IsLzxBMXpfCz0qVN-4mAe673PPpQ/edit?usp=sharing")

```

How to use these functions within this package.

# Cleaning and Organizing the data

The first function in this package is a function that cleans or organizes the data set.This function is called scrap_col() it requires a data set and a column list you wish to get exclude from your data. Don't for to concatenate the columns and use quotations!

Type in this code into your R console.
```{r}
scrap_col(AGP_data, c("recap","notes","username","recent","tagnum"))
```

If the code runs properly then R should return a data set without the specified columns.Limiting the data frame to manageable amounts of columns and rows.

# Manipulating the data by using averages

The next function is going to manipulate the data set by creating a new column and populating it with averages. The function data_mean() uses your own input for a numeric column and takes the mean of the specified column.

Type in this code into your R console.
```{r}
data_mean(AGP_data,mass)
```

If the function works properly, you will see a new column toward the end of the data frame called "mean_column" and it will be populated with averages of the numbers in the column specified.

# Analyzing linear relationships
The next function is for linear models. Using the lm_model() function we can look at the significance and the relationships between the data.

Type in this code into your R console.
```{r}
lm_model(AGP_data,"svl","stage","critter")

```

If the code works properly then you should see a summary table indicating significant relationships between the response and predictor variables of the data set.

# Visualizing the Data using plots
The next function will help visualize data using a variety of plot. Use the custom_plot() function and specify which plot type you wish to see for your data using the independent and dependent variable you chose.

Type in this code into your R console.
```{r}
custom_plot(AGP_data,stage,svl,critter,"scatter")

```

If the function properly works the specified plot type should show and the columns appear in the order you specified in the argument of the function.

# MANOVA analysis
The next function is a MANOVA analysis which uses two numeric response variable and a single categorical predictor. Using the manova_maker() and the data frame and three different columns meeting the criteria above runs a linear model summary similar to a normal linear model function.

Type in this code into your R console.
```{r}
manova_maker(AGP_data,"mass","svl","critter")

```

If the MANOVA function works the result shows the significance of the relationship and uses the Pillai value to show the R\^2 value used in statistical analyses.

# Density Plot

This function creates a density plot using a numeric column and two categorical columns of the data set. Use density_plot() for looking at the fluctuation of the numerical values.

Type in this code into your R console.
```{r}
density_plot(AGP_data,"mass","critter","stage")

```

If the function works properly in R the density plot should appear using the specified columns meeting the aforementioned criteria.
